Chaotic diffusion of tidal streams

### 2015-05-28 ms.tex
writing writing writing

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}
\newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\ident}{\mathbb{1}}
\newcommand{\inttime}{t_{\rm int}}
\newcommand{\fdrate}{\mathcal{R}_\Omega}

@@

The dark matter halos of galaxies are thought to be triaxial in shape. Despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified. Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection. From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii. The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from in-falling satellite galaxies).

As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential. If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter known as stellar tidal streams. The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and mass distribution of the host galaxy. Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies (cite many). Around the Milky Way, streams are found over a large range of distances ($\approx$10-100~kpc). The morphological evolution of the debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also set by the shape and radial profile of the gravitational potential of the host galaxy. By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution around the Milky Way.

Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution. Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}. All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms. Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but claim to extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).

One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} models \citep{kuepper12}. They use the observed SDSS number density of stars and a limited number of radial velocities measured for stream members to fit model streams to the data. In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream. In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster. % In order to extend tidal stream models to more generic potentials, it is important to understand why this stream-fanning occurs and whether this and other strange debris morphologies are generic to more complex potentials.

The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the oblate potential. It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure in the potential. Potentials (Hamiltonians) with more than two degrees of freedom generically contain significant chaotic regions. \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent but found that the orbit is consistent with being regular over dynamically relevant timescales (many Hubble times). It has been shown previously that along some strongly chaotic orbits, tidal streams form large, diffuse fans of debris \citep[e.g.,][]{fardal14}. However, it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity. The result from \citet{pearson15} suggests that even a weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.

TODO: this paragraph needs to be reworked...
In this work, we analyze whether the final configuration-space morphology of simple tidal stream models is correlated with measures of chaos local to the orbit of the progenitor. The effect of chaos in the halos of galaxies has largely been neglected because the chaotic timescales estimated were thought to be many times the age of the universe, but recent work has shown that even weak chaos may be relevant for tidal stream morphology. In this Article, we choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos for grids of constant-energy orbits as computed from single-orbit diagnostics, and compare these results to measures of the mixing time of finite-volume ensembles of orbits (analogs to tidal debris). Finally, we connect the ensemble mixing rate to observational properties of streams, thus demonstrating that local orbital measures of chaos are predictors of tidal stream morphology. This provides a strong case for developing a method to use the mere existence of the many thin streams found in the halo of our own Galaxy to rule out classes or shapes of possible potentials that fit the Milky Way.

This paper is organized as follows: We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}. In Section~\ref{sec:methods}, we describe our choice of potential and method for numerical orbit integration. We present our results in Section~\ref{sec:results} and discuss their implications in Section~\ref{sec:discussion}. We conclude in Section~\ref{sec:conclusions}.

% , we describe the connection between different measures of chaotic behavior, further discuss the implications of chaotic mixing, and comment on the applicability of the methods presented here to more realistic Galactic potentials and simulations.

%[Does this go anywhere?]
%Thin stellar streams form when the spreading of the debris is faster in some preferred direction \citep[one eigenvalue of the Hessian of the potential is large relative to others e.g.,][]{helmi99, sanders??}. This is generically the case for orbits with significant angular momentum in a spherical potential ($|L| > J_r$, orbits that we expect for form streams), but [... comments about expectations in more complex potentials] In a spherical or mildly axisymmetric potential, the debris can only really spread in a single dimension because somehow the constraint that .... Along chaotic orbits, Hessian not defined and frequencies diffuse, ...

@@

% TODO: move all except chaotic diffusion stuff to appendix?

An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series, 
\begin{equation}
(w_1(t),...,w_{2N}(t)) = (q_1(t),...,q_{N}(t),p_1(t),...,p_{N}(t)) \label{eq:coords}
\end{equation}
where $q_i$ and $p_i$ are conjugate coordinates in the sense that
\begin{align}
	\dot{p}_i &= -\frac{\partial H}{\partial q_i}\\
	\dot{q}_i &= \frac{\partial H}{\partial p_i}
\end{align}
for all $t$. The motion in any component, $w_i(t)$, can be represented as a Fourier sum,
\begin{equation}
	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t}; \, (a_k \in \mathbb{C}). \label{eq:fourier}
\end{equation}

A \emph{regular orbit} is a set of such time series that can be transformed to a special set of canonical coordinates known as angle-action coordinates. In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$. The frequencies in the Fourier sum of any individual component of motion (the $\omega_k$ in Equation~1) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
\begin{align}
	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega}; \, (\boldsymbol{n}_k \in \mathbb{Z}^3). \label{eq:fourierfreq}
\end{align}
The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion. Even stronger, the actions are isolating integrals and any pair are in involution such that
\begin{equation}
	[J_i, J_j] = 0
\end{equation}
where $[\cdot,\cdot]$ is the Poisson bracket. This implies that for an $N$ dof system, a regular orbit has $N$ independent constants of motion and the motion is therefore restricted to an $N$-dimensional manifold embedded in the 2$N$ dimensional phase space. The conjugate angle variables define an orthogonal basis, and the topology of angle-action space is therefore the product of $N$ circles, an $N$-torus. Any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus. Each set of actions, $(J_1,...,J_N)$, uniquely labels a torus, and regular orbits may be referred to in terms of their orbital tori. If a transformation to angle-action variables exists, the Hamiltonian may be written solely in terms of the actions,
\begin{equation}
	H = H(\boldsymbol{J})
\end{equation}

A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- that is, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position. The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}.

Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials. Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin56, kuzmin73, deZeeuw85, cincotta06}. The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits. Unlike the box orbits, the tube orbits preserve a sense of rotation about either the long or short axis.

Over long times, a generic, \emph{conditionally-periodic} orbit in this potential fills the surface of its 3-torus --- that is, the three fundamental frequencies of the orbit are incommensurable. If there exists a relation of the form
\begin{equation}
	\boldsymbol{n} \cdot \boldsymbol{\Omega} = n\, \Omega_1 + m\, \Omega_2 + l\, \Omega_3 = 0
\end{equation}
then the orbit is a resonant orbit and the effective dimensionality of the orbit is decreased to 2D. We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits. If an additional resonance relation exists, the orbit is closed and becomes a 1D curve; we refer to such orbits as \emph{bi-resonant} orbits. The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential. In frequency-space, the resonances are planes, but in action-space are generic 2D surfaces. % \cite{cincotta06} analytically compute the resonant structure in action-space for a St\"ackel potential.

The orbit structure of non-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
\begin{equation}
	H(\boldsymbol{J}, \boldsymbol{\theta}) = H_0(\boldsymbol{J}) + \epsilon \, H_1(\boldsymbol{J}, \boldsymbol{\theta})
\end{equation}
where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}. When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by chaotic layers where stochastic motion occurs. Stochastic or chaotic orbits are characterized by apparently random behavior despite being generated by entirely deterministic systems. The stochastic layers form gaps in action space where tori are destroyed. The orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies evolve with time. For small values of $\epsilon$, many regular orbits survive and only small chaotic islands are introduced. As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed. Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.

When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}. We therefore expect a large number of regular orbits will survive --- the so-called KAM tori \citep{} --- however the tori that survive will be separated by regions of chaotic motion. Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion. 
%
%Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials. Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.

@@

The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}. With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, recent NiHao stuff}. It is less clear what happens to the outer halo where tidal streams are more readily found. %; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.

 \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the classic NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos. JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{??, NiHao} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}. JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).

@@

This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies. Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors (cite), radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11, NiHao}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.][]{bailin05}. We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.

@@

\subsection{Numerical approximation of fundamental frequencies}\label{sec:naff}

Bounded, regular orbits in a triaxial potential will have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hamming-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$). It has been shown that using a filter makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time. Once the strongest frequencies of the time series are found, the spectrum must be searched for the three fundamental frequencies.

This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep{} and galaxy dynamics \citep{valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested our own open-source implementation of NAFF in \texttt{Python}.\footnote{This code is publicly available through the \texttt{Gary} package, \url{http://...}} Though our implementation differs slightly from the original definition and from that used in \cite{valluri98}, we still refer to the method as NAFF. For more details about NAFF, see Appendix~\ref{sec:naffapdx} or \cite{papaphilippou96, laskar, etc.}.

>>>>

% commands
\newcommand{\given}{\,|\,}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\transpose}[1]{{#1}^{\mathsf{T}}}
\newcommand{\inverse}[1]{{#1}^{-1}}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\msun}{\ensuremath{\mathrm{M}_\odot}}
\newcommand{\bs}[1]{\boldsymbol{#1}}
\newcommand{\ident}{\mathbb{1}}
\newcommand{\inttime}{t_{\rm int}}
%\newcommand{\fdrate}{\mathcal{R}_\Omega}

@@

The dark matter halos of galaxies are thought to be triaxial in shape. However, despite suggestive evidence from a range of complimentary observational methods, this fundamental prediction from $\Lambda$CDM cosmology has not been conclusively verified. Around other galaxies, it is generally hard to measure the 3D mass profile because they are seen in projection. From the Earth's position within the Milky Way, our view of our own halo and proximity gives us a unique chance to directly measure the 6D positions of stars and model the shape of the mass distribution at large radii. The Milky Way halo has a low density of visible tracers, but luckily many of the halo stars are likely on non-random orbits and thus contain extra information (e.g., debris stripped from stellar systems).

As a satellite galaxy or globular cluster orbits within some larger system, mass is eroded due to the tidal forces of the host galaxy potential. If the orbit of such an object is regular, only mildly eccentric, and stays far from sharp density features (e.g., the disk), the disruption is a steady process and the mass is slowly stretched into long streams of matter known as stellar tidal streams. The phase-space density and therefore the observable properties of these tidal streams are sensitive to the internal progenitor properties, orbit of the progenitor, and mass distribution of the host galaxy. Many tidal streams are observed around the Milky Way, M31, and other nearby galaxies that span a large range of distances \citep[$\approx$10--100~kpc][]{cite many}. The morphological evolution of tidal debris depends on the spread of orbital properties (e.g., actions or frequencies) of the debris and the orbit of the progenitor system, both of which are also determined by the shape and radial profile of the gravitational potential of the host galaxy. By simultaneously modeling the observed phase-space density of stream stars along with the potential, it is hoped that we may infer the dark matter distribution around the Milky Way.

Tidal streams are generally modeled as 1D structures formed on regular orbits with simple internal progenitor dynamics and density evolution. Methods span a range of complexity from orbit-fitting, to \emph{Streakline} or particle-spray models, to action-space density modeling, to N-body simulations \citep[see, e.g., the introduction of][]{kuepper15}. All methods have been tested in some way on simulated observations of data and these tests typically demonstrate the recovery of parameters for analytic, static potential forms. Many of the methods have been shown to work in axisymmetric or otherwise simple potentials but may extend to more complex, generalized potentials (e.g., triaxial or multi-component potentials).

One example of stream modeling in a multi-component (static, analytic) potential was done by \citet{pearson15}, who tried to reproduce observations of the stellar stream density from the globular cluster Palomar 5 in a single oblate and single triaxial potential using \emph{Streakline} \citep{kuepper12} and N-body models. They use the observed SDSS number density of stream stars, a limited number of radial velocities for stream members, and the sky position, distance, radial velocity, and proper motion of the cluster itself to fit model streams to the data. In the oblate potential (a three-component bulge+disk+spherical halo potential), a thin model stream is easily found that reproduces the observed stellar density morphology of the stream. In the triaxial potential (the potential from \cite{law10}: a three component bulge+disk+triaxial halo fit to Sagittarius stream data), the model streams generically form large, two-dimensional ``fans'' of debris near the ends, and there are no physically reasonable progenitor orbits that reproduce the observed thinness of the stream given the observational constraints of the present-day position and velocity of the cluster.

The obvious difference between the two potentials considered by \citet{pearson15} is the extra symmetry of the triaxial potential. It is well known that the number of degrees of freedom of a potential plays a critical role in determining the orbit structure of the potential --- Hamiltonians with more than two degrees of freedom generically contain significant chaotic regions. \citet{pearson15} tested the stochasticity of the orbit of the progenitor that produced fanned debris by computing the Lyapunov exponent along this orbit but found that it is consistent with being regular over dynamically relevant timescales (many Hubble times). It has been shown previously that along some strongly chaotic orbits, tidal streams do form large, diffuse fans of debris \citep[e.g.,][]{fardal14}, however it is unknown how the resultant properties of the debris (e.g., density or length of the stream) depend on the degree of stochasticity. The result from \citet{pearson15} suggests that even weak chaos (as measured by the Lyapunov exponent) may affect the density evolution and therefore observability of tidal streams.

In this work, we study the effect of chaotic diffusion on tidal stream morphology. We choose a simple, cosmologically motivated model for a triaxial potential, analyze the degree of chaos as computed from single-orbit diagnostics for grids of constant-energy orbits, and compare these results to measures of the density evolution of finite-volume ensembles of orbits (meant to mimic tidal debris). We find that even when the chaotic timescale is predicted to be long for a given orbit, chaos may manifest over much shorter times in small orbit ensembles due to the nature of chaotic diffusion. This result supports a reevaluation of the importance of chaos in galactic halos and implies that the amount of chaos in a given potential may have significant consequences for the observability and survivability of thin, cold tidal streams.

This paper is organized as follows. We review relevant nonlinear dynamics in Section~\ref{sec:nldreview}. In Section~\ref{sec:methods}, we describe our choice of potential, method for numerical orbit integration, and introduce the chaos indicators used in this work. Our results are split into three subsections: in Section~\ref{sec:results1} we present iso-energy grids of orbits and discuss the orbit classes and chaotic timescales present in our potential; in Section~\ref{sec:results2} we study the density evolution of small ensembles of orbits around each orbit of the previous section; and, in Section~\ref{sec:results3} we describe the behavior of chaotic diffusion and use this to explain how chaos is relevant for tidal streams over short times. We discuss the implications of our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.

@@

An orbit in an $N$ degree of freedom (dof) Hamiltonian, $H$, is a set of $2N$ quasi-periodic time series, 
\begin{equation}
(w_1(t),...,w_{2N}(t)) = (q_1(t),...,q_{N}(t),p_1(t),...,p_{N}(t)) \label{eq:coords}
\end{equation}
where $q_i$ and $p_i$ are conjugate coordinates in the sense that
\begin{align}
	\dot{p}_i &= -\frac{\partial H}{\partial q_i}\\
	\dot{q}_i &= \frac{\partial H}{\partial p_i}
\end{align}
for all $t$. If bounded, the motion in any component, $w_i(t)$, can be represented as a Fourier sum,
\begin{equation}
	w_i(t) = \sum_k^\infty a_k \, e^{i\,\omega_k\,t} \label{eq:fourier}
\end{equation}
where the $a_k$ are complex amplitudes.

A \emph{regular orbit} is a set of such time series that can be transformed to a special set of canonical coordinates known as angle-action coordinates. In these coordinates, the position variables are angles, $\boldsymbol{\theta}$, that increase linearly with time with rates set by $N$ constant, fundamental frequencies, $\boldsymbol{\Omega} = (\Omega_1, ..., \Omega_N)$. The frequency of a Fourier component in the sum for any individual component of motion (the $\omega_k$ in Equation~\ref{eq:fourier}) are just linear, integer combinations of the fundamental frequencies, $\boldsymbol{\Omega}$ --- that is, for a regular orbit, any Fourier component frequency may be written
\begin{align}
	\omega_k &= \boldsymbol{n_k} \cdot \boldsymbol{\Omega} \label{eq:fourierfreq}
\end{align} %; \, (\boldsymbol{n}_k \in \mathbb{Z}^3)
where $\boldsymbol{n}_k$ is a vector of $N$ integers. The conjugate momentum coordinates --- the actions,  $\boldsymbol{J}$ --- are constants of motion. Even stronger, the actions are isolating integrals and any pair are in involution such that
\begin{equation}
	[J_i, J_j] = 0
\end{equation}
where $[\cdot,\cdot]$ is the Poisson bracket. This implies that for an $N$ dof system, a regular orbit has $N$ independent constants of motion and the motion is therefore restricted to an $N$-dimensional manifold embedded in the 2$N$ dimensional phase space. In fact, the topology of angle-action space is toroidal and any regular orbit in an $N$ dof Hamiltonian can be understood as motion on the surface of an $N$-torus. Each set of actions, $(J_1,...,J_N)$ (or frequencies), uniquely labels a torus, and regular orbits are sometimes referred to in terms of their orbital tori.

A Hamiltonian or potential is said to be \emph{globally integrable} when the number of isolating integrals of motion is equal to the number of degrees of freedom and a transformation to angle-action coordinates may be done globally --- for example, the transformation to angle-action coordinates may be written as a function of arbitrary phase-space coordinates and the functional form is independent of phase-space position. The condition for global integrability is very restrictive and it seems that the likelihood that a Hamiltonian is globally integrable decreases as the number of degrees of freedom increase \citep[e.g.,][]{lichtenberg83}. In a globally integrable potential, the Hamiltonian may be written solely in terms of the actions, $H = H(\boldsymbol{J})$. Galactic potentials are almost certainly not globally integrable, but it is useful to understand the orbit structure in integrable systems before extending to more general potentials. Only one globally integrable triaxial mass model is known that generates motion that is qualitatively similar to that seen in Galaxies: the St\"ackel or ``Perfect Ellipsoid'' model \citep[e.g.,][]{kuzmin73, deZeeuw85}. The four general classes of orbits in this potential are box, inner long-axis tube, outer long-axis tube, and short-axis tube orbits. The tube orbits preserve a sense of rotation about either the long or short axis and are therefore centrophobic, whereas the box orbits may approach arbitrarily close to the center of the potential and are centrophilic.

The frequencies of a generic, non-resonant orbit are incommensurable --- that is, $\bs{n} \cdot \bs{\Omega} \neq 0$ for any integer vector, $\bs{n}$, with reasonable magnitude.\footnote{A more precise condition is stated in terms of a diophantine condition, e.g., $|\bs{n} \cdot \boldsymbol{\Omega}| > \alpha \, |n|^{-\gamma}$ where $\alpha, \gamma>0$.} Conditionally-periodic orbits uniformly cover the surface of their orbital torus. If instead there exists a relation of the form $\boldsymbol{n} \cdot \boldsymbol{\Omega} = 0$ then the orbit is a resonant orbit. For resonant orbits, the existence of an additional \emph{local} constant of motion reduces the dimensionality of the orbit. We refer to orbits that obey a single resonance relation as \emph{uni-resonant} orbits, and if an additional resonance relation exists, \emph{bi-resonant}. The resonant structure of a potential --- the relative importance of particular resonance numbers --- is difficult to compute, but determines the global behavior of orbits in the potential. For a triaxial potential the resonances define planes in frequency-space, but resonance surfaces in action-space are generic 2D surfaces.

The orbit structure of non-integrable potentials can be understood by considering a Hamiltonian that is a small perturbation away from being globally integrable --- that is, a Hamiltonian that may be written
\begin{equation}
	H(\boldsymbol{J}, \boldsymbol{\theta}) = H_0(\boldsymbol{J}) + \epsilon \, H_1(\boldsymbol{J}, \boldsymbol{\theta})
\end{equation}
where $H_0(\bs{J})$ represents an integrable Hamiltonian, and $\epsilon$ is a small parameter that determines the perturbation strength \citep[a complete description of perturbation theory applied to nonlinear Hamiltonians is given in][]{lichtenberg83}. When $0 < |\epsilon| \ll 1$, resonant surfaces become ``thick'' resonant layers, within which orbits are qualitatively similar to the parent resonant orbit \citep[e.g.,][]{merritt99} and the stable regions are surrounded by stochastic layers where chaotic motion occurs. Chaotic orbits are characterized by exhibiting random behavior despite being generated by entirely deterministic equations of motion. Stochastic layers form gaps in action space where tori are destroyed; the orbits that exist in these regions therefore can not be represented in angle-action coordinates, as their actions and frequencies are not constant and diffuse chaotically with time. For small values of $\epsilon$, many regular orbits survive and only small chaotic islands are introduced. As the strength of the perturbation increases, eventually all tori associated with conditionally-periodic motion will be destroyed. Next, the uni-resonant tori are destroyed, and finally the bi-resonant tori --- these are least susceptible to destruction from perturbations \citep[e.g., see][]{valluri98}.

When $\epsilon$ is large, there is no general prediction for the resulting behavior, however it seems that more complicated and physically motivated triaxial potential models for galaxies follow the intuition gained from the small-perturbation picture described above, at least for certain parameter choices \citep{valluri, merritt, dehnen, etc.}. We therefore expect a large number of regular orbits will survive --- the so-called KAM tori --- however the tori that survive will be separated by regions of chaotic motion. Any transformations to angle-action coordinates must be defined locally due to the destruction of tori caused by the onset of chaotic motion. [... how and where to end this section?]
%
%Analytic transformations to angle-action coordinates are only known for a handful of simple, integrable potentials. Much effort has gone in to developing general methods for transforming either from position-velocity to angle-action coordinates, $(\bs{x},\bs{v})\rightarrow(\bs{\theta},\bs{J})$ \citep{many}, or vice-versa, $(\bs{\theta},\bs{J})\rightarrow(\bs{x},\bs{v})$ \citep{many}.

@@

The density distributions within dark matter halos formed in cosmological N-body simulations are generically triaxial \citep[e.g.,][]{jing02, bett07, zemp09, veraciro11}. With the inclusion of baryonic physics and sub-grid prescriptions for energy input due to supernovae and other feedback mechanisms, the inner potential ($\lesssim$$0.1R_{\rm vir}$ for a $\approx$$10^{12}~\msun$ halo mass) typically becomes more spherical, though the magnitude of this reshaping depends on the particular merger history and star formation efficiency within a given halo \citep[e.g.,][]{dubinski??, butsky15}. It is less clear what happens to the outer halo where tidal streams are more readily found. %; for this reason, we ask what we would expect to happen to tidal streams if the halo in which they form is strongly triaxial.

 \citet[][hereafter JS02]{jing02} found that a triaxial generalization of the NFW density profile \citep{navarro96} generates excellent fits to the density distributions within haloes in their high-resolution N-body simulations, and they provide probability distributions for the axis ratios of a large sample of these halos. JS02 finds median axis ratios of $c/a \approx 0.55$ and $b/a \approx 0.77$ where $a$ is the major axis, $b$ the intermediate, and $c$ the minor axis.\footnote{Note that JS02 use the opposite notation so that $c$ is the major and $a$ is the minor axis.} These are largely consistent with findings from more recent simulations \citep[e.g.,][]{millenium, hmm, butsky15} and consistent with constraints from weak lensing that place a lower limit on minor-to-major axis ratios of $c/a\gtrsim0.5$ \citep{vanuitert12}. JS02 finds significant scatter in the distributions of concentration parameter, $c_e$, or scale radius (depending on choice of parametrization).

@@

This potential is a simple and unrealistic model for the total potential of a Milky-Way-like galaxy, however it represents a conservative choice for exploring the structure of orbits in the halos of such galaxies. Realistic galactic potentials will have a significant component due to the disk and bulge, may have twisting inertia tensors \citep{romanowsky98}, radially changing axis ratios \citep[e.g.,][]{kazantzidis04,debattista08,veraciro11,butsky15}, significant substructure \citep{moore98,zemp09}, or time dependence \citep[either from bulk rotation, mass growth, mergers, etc.; see, e.g.,][]{bailin05}. We expect inclusion of any of these effects to increase the amount and influence of chaos; see the Discussion (Section~\ref{sec:discussion}) for a few simple demonstrations.

@@

\subsection{Frequency diffusion rate}\label{sec:naff}

Bounded, regular orbits in a triaxial potential have three fundamental frequencies, $\bs{\Omega}$, that determine the periodic behavior of motion. The motion in any canonical coordinate can therefore be decomposed as a Fourier sum (Equation~\ref{eq:fourier}) where the Fourier frequencies are linear, integer combinations of the fundamental frequencies (Equation~\ref{eq:fourierfreq}). \cite{laskar93} introduced a method for recovering the fundamental frequencies of an orbit that effectively uses a Hamming-filtered fast-Fourier transform (FFT) of complex combinations of the motion (e.g., $x(t) + i v_x(t)$). It has been shown that using a filtered FFT makes the accuracy in determining the frequencies converge much faster than the typical $\inttime^{-1}$ expected for a standard FFT \citep{laskar99}, where $\inttime$ is the integration time. The FFT is iteratively searched for the strongest frequency component, which is subtracted from the spectrum after each iteration. This procedure generates a table of frequencies for each component of motion which must then be searched for the three fundamental frequencies.

This method is referred to as ``Numerical Approximation of Fundamental Frequencies'' (NAFF) and has been used extensively in planetary dynamics \citep[e.g.,][]{laskar93b, laskar96} and galaxy dynamics \citep{papaphilippou98, valluri98}, especially in the study of orbits in triaxial systems. We have implemented and tested our own open-source implementation of NAFF in \texttt{Python}.\footnote{This code is publicly available through the \texttt{Gary} package, \url{http://...}} Though our implementation differs slightly from the original definition and from that used in \cite{valluri98}, we still refer to the method as NAFF. For more details about NAFF, see Appendix~\ref{sec:naffapdx} or, e.g., \cite{laskar88, laskar93, papaphilippou96}.
